\documentclass[twocolumn,DIV20]{scrartcl}
\usepackage{units}
\usepackage{verbatim}
\title{Simple Raytracer in Common Lisp}
\author{Martin Kielhorn}
\date{2010-09-05}
\begin{document}
\maketitle
\section{Introduction}
We design a device consisting of multiple light diodes that are fired
in sequence. LEDs can deliver more light when run in a pulsed mode. So
in principle it should be possible to obtain higher brightness inside
the field of view of a microobjective.

Our optical setup consists of several lenses to reimage the light
emitting area into the back focal plane of the objective and one
movable mirror. The light emitted from the diode is assumed to be
lambert distributed. In order to find good focal length ratios and the
diameters of the lenses the optical system is simulated by raytracing
(neglecting aberrations).

The raytracer consists of several distinct modules whose function and
implementation will be described in this document.

\section{Vector}
The fundamental datastructure are three-dimensional vectors. In the
program they are represented as an array with three elements. The type
of the elements (called vec-float) can be set to either single-float
or double-float. For the simulation the latter choice is
endoresed. 
\subsection{Single or double-float?}
However, some effort has been put into making it possible
to switch to single-float at compile time. This effort consists mainly
of replacing numerical constants like 1d0 with the constant +one+ that
is automatically set to 1s0 or 1d0 depending on the type vec-float.

A three-dimensional vector can be created with make-vec. This expects
zero to three arguments and defaults to zero. If all arguments are
constant numerals a macrocall like \verb!(v 0 2.3)! should be
used. It coerces the arguments to be of the right type for vec-float,
obliviating cumbersome calls like \verb!(make-vec :y 2.3d0)!
(which would only work for double-float) or the even more cumbersome
notation \verb!(make-vec :y #.(* +one+ 2.3d0))!
which would enforce conversion to double-float at compile time if
necessary.

The main motivation to introduce the extra type vec-float is error
estimation. If running a simulation with single-float gives the same
result as with double-float we can be quite sure that round-off errors
are not skewing our result.

\subsection{Operations on vectors}
Addition v+, subtraction v-, multiplication with scalar (v* scalar v),
dot product v. and cross product vx are working as expected. The
euclidean length of a vector is given by norm and the result of the
function normalize is a unit vector with the same direction as the
1input vector.

\subsection{Matrix operations}
The function m takes 9 parameters and creates a new $3\times3$
matrix. The function \verb!(rotation-matrix angle v)! takes an angle
in radians and a unit vector and defines the corresponding rotation
around an axis. The function \verb!(m* m v)! applies a matrix to a
vector.

\subsection{Macros}
Sometimes the macro with-arrays (found in Nikodemus' raylisp) is
handy. It obliviates the need to write aref for function accesses.

\section{Graphical User Interface}
The macro with-gui opens a window calling the body repeatedly to draw
the contents. It makes sense to call a function draw inside
with-gui. Everytime draw is updated from within slime the changes are
immediately visible on the screen.

\section{Optics}
This module contains the basic raytracer.
\subsection{Objects}
The objects plane, mirror, lens, objective and ray are implemented as
classes. A plane is defined by a contained point (called center) and a
normal. A mirror additionally contains a radius and a lens its
lens-radius and focal-length. An objective is a subclass of length
augmented with immersion-index, numerical-aperture and bfp-radius. The
lens-radius of an objective isn't important and should be set to 10
times bfp-radius. A ray consists of a start position and a direction
vector. In general the direction of a ray should have length one.
\subsection{Raytrace}
The generic methods intersect, refract and reflect take a ray and an
object as parameters and return a new ray. If the ray doesn't pass
through an aperture or hits an objective in a too steep angle the
condition ray-lost is signalled. All methods check that directions are
unit vectors. Note that the resulting direction of raytrace on a lens
isn't normalized.
\subsection{Light source}
The function lambert-direction returns a normalized direction vector
with Lambert-Cosine $\theta$ distribution (elevation) and uniform
$\phi$ distribution (azimuth).
\subsection{Helpers}
The function back-focal-plane-radius returns the radius of the back
focal plane of a high NA immersion objective. The function
focal-length-from-magnification can be used to calculate the focal
length from the given magnification for Zeiss objectives. Both of
these functions are called from make-objective, which will construct
an immersion objective from a given NA, magnification and immersion
index.

The functions grad->rad and rad->grad can be used to convert degree
into radians and vice versa.

The function etendue calculates the optical invariant if chief and
marginal ray are known. More useful for our purposes is the more
specific oil-objective-etendue that calculates the optical invariant
necessary to illuminate a radius in the focal plane of an immersion
objective.

The function magnification-from-angles can be used to find the
magnification of an imaging system given the marginal ray angles on
object and image side.

\section{Specific functions}
This module contains specific functions that are needed to construct
the optical imaging path through the illuminator. The function
led-half-angle-to-achieve-etendue calculates the angle that needs to
be captured to fill the full field of view in an objective.

The function led-magnification-to-fit-in-galvo-interval says how much
the LED angle has to be reduced to fit enough of them inside of the
range that the movable mirror can cover.


\end{document}